Is high social class always beneficial for survival? Northern Sweden 1801--2013

### 2018-01-31 particle3.Rmd
fixes

## Ages 40--64

```{r getske4060}
cap1w <-  "Cumulative hazards for social class by time period, ages 40--64, women."
scap1w <-  "Cumulative hazards, ages 40--64, women."
cap1m <-  "Cumulative hazards for social class by time period, ages 40--64, men."
scap1m <-  "Cumulative hazards, ages 40--64, men."
##
cap0m <- "Median age at death for a 40-year-old by social class and period, women (top) and men (bottom)."
scap0m <- "Median age at death for a 40-year old by social class."

@@

```{r hazcalw}
source("R/pgetPHglm2.R")
source("R/getPHglm.R")
##nper <- 4
nper <- length(labb)
resw <- pgetPHglm2(vb40[vb40$sex == "female", ],
            nper, labb)
resm <- pgetPHglm2(vb40[vb40$sex == "male", ],
            nper, labb)
```

```{r meli, fig.cap = cap0m, fig.height = 8, out.height = "14cm"}
oldpar <- par(mfrow = c(2, 1))
plot(1:13, resw[1, ], type = "b", col = scol[1], axes = FALSE, ylim = c(40, 75), main = "Women", ylab = "Age", xlab = "", xlim = c(1, 15.5))
axis(1, at = 1:13, labels = labb, las = 2)
axis(2, at = c(40, 50, 60, 70), las = 1)
abline(h = c(50, 60, 70), lty = 3)
box()
lines(1:13, resw[2, ], type = "b", lty = 2, pch = 2, col = scol[2])
lines(1:13, resw[3, ], type = "b", lty = 3, pch = 3, col = scol[3])
lines(1:13, resw[4, ], type = "b", lty = 4, pch = 4, col = scol[4])
abline(h = 40)
legend("bottomright", legend = c("elite", "middle", "worker", "NA"), lty = 1:4, pch = 1:4, col = scol)
## Men:
plot(1:13, resm[1, ], type = "b", col = scol[1], axes = FALSE, ylim = c(40, 75), main = "Men", ylab = "Age", xlab = "", xlim = c(1, 15.5))
axis(1, at = 1:13, labels = labb, las = 2)
axis(2, at = c(40, 50, 60, 70), las = 1)
box()
lines(1:13, resm[2, ], type = "b", lty = 2, pch = 2, col = scol[2])
lines(1:13, resm[3, ], type = "b", lty = 3, pch = 3, col = scol[3])
lines(1:13, resm[4, ], type = "b", lty = 4, pch = 4, col = scol[4])
abline(h = 40)
abline(h = c(50, 60, 70), lty = 3)
legend("bottomright", legend = c("elite", "middle", "worker", "NA"), lty = 1:4, pch = 1:4, col = scol)
par(oldpar)
source("R/plotED.glm.R") # From below (not eval)
```

```{r cumsw, fig.cap = cap0, fig.scap = scap0, fig.height = 8, out.height = "14cm",eval = FALSE}
source("R/plotED.glm.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = TRUE, ylab = "H(65|40)",
                main = "Women 40-64", ylim = c(0, 1))##, exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = TRUE, ylab = "H(65|40)",
                main = "Men 40-64", ylim = c(0, 1))##, exclude = 5:6)
par(oldpar)
```

>>>>

## Median remaining life at age 40

```{r getske4060}
cap1w <-  "Cumulative hazards for social class by time period, ages 40--64, women."
scap1w <-  "Cumulative hazards, ages 40--64, women."
cap1m <-  "Cumulative hazards for social class by time period, ages 40--64, men."
scap1m <-  "Cumulative hazards, ages 40--64, men."
##
cap0m <- "Median age at death for a 40-year-old by social class and period, women (top) and men (bottom)."
scap0m <- "Median age at death for a 40-year old by social class."
cap0md <- "Differences in median age at death for a 40-year-old by social class and period (ref: middle class), women (top) and men (bottom)."
scap0md <- "Differences in median age at death for a 40-year old by social class."

@@

```{r hazcalw}
source("R/pgetPHglm2.R")
source("R/getPHglm.R")
##nper <- 4
nper <- length(labb)
resw <- pgetPHglm2(vb40[vb40$sex == "female", ],
            nper, labb)$medi
resm <- pgetPHglm2(vb40[vb40$sex == "male", ],
            nper, labb)$medi
```

```{r meli, fig.cap = cap0m, fig.height = 8, out.height = "14cm"}
oldpar <- par(mfrow = c(2, 1))
plot(1:13, resw[1, ], type = "b", col = scol[1], axes = FALSE, ylim = c(60, 95), main = "Women", ylab = "Age", xlab = "", xlim = c(1, 15.5))
axis(1, at = 1:13, labels = labb, las = 2)
axis(2, at = c(60, 70, 80, 90), las = 1)
abline(h = c(60), lty = 3)
abline(h = c(70, 80, 90), lty = 3)
box()
lines(1:13, resw[2, ], type = "b", lty = 2, pch = 2, col = scol[2])
lines(1:13, resw[3, ], type = "b", lty = 3, pch = 3, col = scol[3])
lines(1:13, resw[4, ], type = "b", lty = 4, pch = 4, col = scol[4])
abline(h = 60)
legend("bottomright", legend = c("elite", "middle", "worker", "NA"), lty = 1:4, pch = 1:4, col = scol)
## Men:
plot(1:13, resm[1, ], type = "b", col = scol[1], axes = FALSE, ylim = c(60, 95), main = "Men", ylab = "Age", xlab = "", xlim = c(1, 15.5))
axis(1, at = 1:13, labels = labb, las = 2)
axis(2, at = c(60, 70, 80, 90), las = 1)
box()
lines(1:13, resm[2, ], type = "b", lty = 2, pch = 2, col = scol[2])
lines(1:13, resm[3, ], type = "b", lty = 3, pch = 3, col = scol[3])
lines(1:13, resm[4, ], type = "b", lty = 4, pch = 4, col = scol[4])
abline(h = 60)
abline(h = c(70, 80, 90), lty = 3)
legend("bottomright", legend = c("elite", "middle", "worker", "NA"), lty = 1:4, pch = 1:4, col = scol)
par(oldpar)
source("R/plotED.glm.R") # From below (not eval)
```

In Figure \@ref(fig:melidiff) the difference in median life (Reference: middle class) is shown.

```{r melidiff, fig.cap = cap0md, fig.height = 8, out.height = "14cm"}
oldpar <- par(mfrow = c(2, 1))
for (m in c(1, 3, 4)) resw[m, ] <- resw[m, ] - resw[2, ]
resw[2, ] <- 0
for (m in c(1, 3, 4)) resm[m, ] <- resm[m, ] - resm[2, ]
resm[2, ] <- 0
plot(1:13, resw[1, ], type = "b", col = scol[1], axes = FALSE, main = "Women", ylab = "Years", xlab = "", xlim = c(1, 15.5), ylim = c(-8, 8))
axis(1, at = 1:13, labels = labb, las = 2)
axis(2, at = c(-5, 0, 5), las = 1)
abline(h = c(-5, 5), lty = 3)
box()
lines(1:13, resw[2, ], type = "b", lty = 2, pch = 2, col = scol[2])
lines(1:13, resw[3, ], type = "b", lty = 3, pch = 3, col = scol[3])
lines(1:13, resw[4, ], type = "b", lty = 4, pch = 4, col = scol[4])
abline(h = 60)
legend("bottomright", legend = c("elite", "middle", "worker", "NA"), lty = 1:4, pch = 1:4, col = scol)
## Men:
plot(1:13, resm[1, ], type = "b", col = scol[1], axes = FALSE, main = "Men", ylab = "Years", xlab = "", xlim = c(1, 15.5), ylim = c(-8, 8))
axis(1, at = 1:13, labels = labb, las = 2)
axis(2, at = c(-5, 0, 5), las = 1)
box()
lines(1:13, resm[2, ], type = "b", lty = 2, pch = 2, col = scol[2])
lines(1:13, resm[3, ], type = "b", lty = 3, pch = 3, col = scol[3])
lines(1:13, resm[4, ], type = "b", lty = 4, pch = 4, col = scol[4])
abline(h = c(-5, 5), lty = 3)
legend("bottomright", legend = c("elite", "middle", "worker", "NA"), lty = 1:4, pch = 1:4, col = scol)
par(oldpar)
source("R/plotED.glm.R") # From below (not eval)
```
```{r cumsw, fig.cap = cap0, fig.scap = scap0, fig.height = 8, out.height = "14cm",eval = FALSE}
source("R/plotED.glm.R")
oldpar <- par(mfrow = c(2, 1))
woutp <- plotED.glm(resw, col = scol, logScale = TRUE, ylab = "H(65|40)",
                main = "Women 40-64", ylim = c(0, 1))##, exclude = 5:6)
moutp <- plotED.glm(resm, col = scol, logScale = TRUE, ylab = "H(65|40)",
                main = "Men 40-64", ylim = c(0, 1))##, exclude = 5:6)
par(oldpar)
```
